---
title: "Using _uspr_"
author: "Martin R. Smith"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
bibliography: ../inst/REFERENCES.bib
csl: ../inst/apa-old-doi-prefix.csl
vignette: >
  %\VignetteIndexEntry{Using uSPR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

_uspr_ is a package to compare phylogenetic trees.

This vignette assumes familiarity with the process of [loading trees into R](https://ms609.github.io/TreeTools/articles/load-trees.html).

To load the `uspr` library into the R namespace, use 

```{r}
library('uspr')
```

We'll work with some simple trees generated using the 
[`TreeTools`](https://ms609.github.io/TreeTools) package.

```{r, align='center', fig.width=7}
library('TreeTools')
tree1 <- BalancedTree(10)
tree2 <- TreeTools::PectinateTree(10)
par(mfrow=1:2, mar=rep(0.2, 4)) # Set up plotting area
plot(tree1)
plot(tree2)
```

To generate an approximate SPR distance between two unrooted trees, use:

```{r}
USPRDist(tree1, tree2)
```

For Replug distances, it's:

```{R}
ReplugDist(tree1, tree2)
```

For TBR distances, use:

```{r}
TBRDist(tree1, tree2, exact = TRUE)
```

When calculating an exact TBR distance, we receive information on the
maximum agreement forest for free (i.e. with no extra processing cost):

```{r}
TBRDist(tree1, tree2, exact = TRUE, maf = TRUE)
```

Once trees have more than about a dozen tips, it becomes slow to calculate the 
exact distance.  In the interests of speed, we may wish to approximate the value
of the TBR distance:

```{r}
TBRDist(tree1, tree2, exact = FALSE)
```

To avoid multiple calls, each function can also be used to compare lists of 
trees (or multiPhylo objects) against a single tree:

```{r}
TBRDist(tree1, list(tree1, tree2), exact = TRUE, maf = TRUE)
```

Or against each corresponding entry in a second list:

```{r}
USPRDist(list(tree1, tree2, tree2), list(tree2, tree1, tree2))
```
